A deeper look into bug fixes: Patterns, replacements, deletions, and additions

### 2016-02-19 MSRChallenge2016.tex
half a column over

There has been considerable attention paid to the problem of automatic
program repair in recent
years~\cite{kim2013,legoues2012,pan2009,Weimer13, Qi13,
  Mechtaev15,prophet}, among many others),
%
One broad class of techniques in this space take a generate-and-validate
approach, which involves generating a large number of candidate patches using a
pre-defined set of mutation operators, and then validating correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

@@

\noindent\textbf{Related work.} A good complement to this paper is~\cite{zhong2015}, which studies 
a similar set of questions in C.  Beyond focusing on Java, we focus more on
replacements, ommitted from that previous analysis, and otherwise target a
larger dataset~\cite{dyer2013}.  Barr et al.~\cite{Barr14fse} studied changes to
Java programs to understand the ``Plastic Surgery Hypothesis'' underlying
certain types of program repair techniques, an important but orthogonal
concern. Kim et al.~\cite{kim2013} manually analyze 
changes to Java programs to inform an automated repair
technique and show that doing so results in higher-quality repairs; 
we attempt to validate their learned templates heuristically on
a different dataset, and take their results as
motivation for understanding human repairs (as it may result in better patches).
Long and Rinard learn probabilistic models from bug fixes to C
code~\cite{Long2016}. Our analysis is significantly less precise and does not
inform a new repair technique.  Rather, it serves as a starting
point for research investigating the repair of bugs in Java code.

@@

\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre-fix and
post-fix version. If there are count differences between pre-fix and post-fix
version, we consider the COB pattern found.

@@

\begin{table*}
	\centering
	\resizebox{\textwidth}{100pt}{
		\begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|} \hline
			&Assert&Break&Continue&Do& For&If&Label&Return&Case&Switch& Synch&Throw&Try&TypeDecl&While
			\\ \hline

@@

After running the bug pattern detection approach on Boa's dataset, we obtained
the result shown in Table~\ref{tab:freqpattern}. It shows that the
most common pattern among these patterns is CBC and the least common pattern is
AOB. If we assume that these patterns never appear together,
they cover 14.78\% of buggy files. Kim et al. mentioned that the
patterns that they use cover almost 30\% of real
patches~\cite{kim2013}. Although we only detect 8 out of 10 bug fixing patterns,
it is hard to believe that the remaining two patterns cover 15\% of buggy files,
especially considering that 14.78\% is a number that we get when we assume that the patterns never appear together. Moreover, the percentage of each pattern is based
on imprecise approximation which is likely to represent an upper bound rather
than a lower bound. This suggests that common bug fixing patterns might not be
as common as it appeared to be.

\section{Statement-level mutations}
Some automatic program repair approaches mutate source code using predefined
mutation operators such as addition, deletion and replacement. In
order to generate patches that make more sense to human developers, it would be
beneficial to analyze how developers mutate source code to fix bugs.

\subsection{Replacements}

Because direct diffs are difficult to identify, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another across
this dataset.  To do this, for each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.  For each statement type (as tagged by the Boa infrastructure
for Java), we process the results as follows: if number of occurrences of one
statement type decreased post-fix and the number of another type increased, we
say that the first statement type was replaced by the second statement type for
that file. It is important to notice that this analysis doesn't distinguish the
replacement of the same statement kind, since we are counting the amount of
appearances of each statement kind.

The numbers in the slots of Table~\ref{tab:likeliness} \todo{shall we add \% in
this table?}, represent the percentage rate at which the elements of the
leftmost column were replaced by the elements of the top row. In the first
column we have the list of statements that are being replaced (faulty statements
that were removed), and on the top row we have the list of statements for which
that statement was replaced for (the new statement put instead). For example, if
we take the intersection of row 5 (DoStatement row) and column 2 (AsserStatement
column) of the data, we have a 0.81, which means that Do statements were
replaced by Assert statements 0.81\% of the times.

Our results show that the most common replacement is replacing Return statement
with If statement, which showed up in 30,489 files. The second most common
replacement is replacing an If statement with a Return statement, which had
28,536 appearances. We found that the most common \emph{Replacer} statement kind
(the statement kind that most commonly replaces others) is the \textbf{If
Statement} with 101,366 appearances; and the least common \emph{Replacer} is the
\textbf{Type Declaration Statement} with 447 appearances.

On the opposite side of the data, the least common replacement was an Assert
statement replacing a Type Declaration statement, which had zero appearances.
The second least common replacement were replacing a Do statement with a Type
Declaration statement and replacing a Label with a Type Declaration, both of
which had one appearance respectively in our search. The most common $Replacee$
(the most likely statement kind to be replaced by another statement kind) is the
\textbf{Return Statement} with 111,938 appearances; and the least common
$Replacee$ is the \textbf{Type Declaration Statement} with 399 appearances.

\subsection{Deletions and Insertions}

@@

For each bug fixing revision $r$ and each statement kind $k$, we count how many
statements of kind $k$ are there in revision $r$ and $r-1$ ($r-1$ means the
revision before revision $r$), and then compare the two numbers.

From Figure~\ref{fig:StmtType}, we can see that, expression, if, return, for and
try catch statements are added most often. Meanwhile, the same types of
statements are also deleted most often. This finding indicates that most bugs
were fixed by changing the control flow of the program.

\subsection{Applicability to Automatic Error Repair}

This study is directly applicable to automatic software repair. Most common
tools for automatic software repair~\cite{kim2013,legoues2012} first
find a faulty statement using several different fault localization
techniques~\cite{fry2010}, and once they have that faulty statement, they apply
different kinds of edits to it, usually at random. Which makes it equally likely
to replace an statement for another without looking at the kinds of statements
being replaced. With this study, we are able to make a more accurate guess based
on the phenomenon we observed from Table~\ref{tab:likeliness} and
Figure~\ref{fig:StmtType}.

For example, if we are fixing a Java program, and we have a faulty statement
that happens to be a For Statement. Then we can look for the most likely
statements to replace a For Statement (row 6): If statement (22.89\%
likeliness), Return statement (21.08\%), and a While statement (13.79\% of
likeliness to replace a For Statement). We also know that there is a 7.9\% of
cases in which human programmers deleted this kind of statement, and in 10.6\%
of the cases, a For statement was added to patch a bug. This way we provide
guidelines to the automatic bug repair approaches to consider this on their
effort to find a patch.

\section{Threats to Validity}

Since we are using a DSL provided by Boa to implement our analysis, the
correctness of our analysis depends on our programs as well as Boa itself. To
prevent errors from our side, we open source our implementation on
Github~\footnote{\url{https://github.com/}}.\todo{update
this}

Some threats to our study are due to the limitation of Boa. For example, we rely
on Boa to identify bug fixing revisions, but it is unclear whether those
identified revisions are actually related to bug fixing. After all, identifying
bug fixing revisions from revision history is still an open problem. Another big
limitation of Boa is that there is no easy way to get a precise statement-level
diff between two revisions. Thus, our template matching and analysis of code
changes (by counting each statement kind) could be inaccurate. However, our
results should provide upper bounds for the questions we investigated.

\section{Conclusions}
We have characterized bug fixing changes in Java.
On average, a bug fixing revision changes 11.3 files and often introduces new
fields and methods. We have also showed that 8 out of 10 bug fixing patterns in
PAR cover only 15\% of bug fixes. Last but not least, we investigated statement
level mutations and summarized the likelihood of statement insertion, deletion,
and replacement.

The findings of our study provide a set of useful insights for automatic
program repair tools in Java. It suggests that patterns in PAR are not enough to
cover the entirety of bug fixes in Java. Adding more patterns is one solution,
but may not be a permanent one. Our finding on Java bug fixes characteristic suggests that mutation based program repair may need to add new fields or methods
to fix bugs in Java and achieve a human-like written patch. A deeper look suggests
that it can also improve its search for a patch by having the knowledge of how
likely it is to replace a faulty statement by the different kinds of statements
available. Overall, our findings demonstrate that our knowledge of bug fixing
properties in Java is still lacking and therefore more research on it are
warranted.

>>>>

There has been considerable attention paid to the problem of automatic
program repair in recent
years~\cite{kim2013,legoues2012,pan2009,
  Mechtaev15,prophet}, among many others),
%
One broad class of techniques in this space take a generate-and-validate
approach, which involves generating a large number of candidate patches using a
pre-defined set of mutation operators, and then validating correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

@@

\noindent\textbf{Related work.} Zhong and Su~\cite{zhong2015} study
a similar set of questions in C.  Beyond focusing on Java, we focus more on
replacements, ommitted from that previous analysis, and target a
larger dataset~\cite{dyer2013}.  Barr et al.~\cite{Barr14fse} studied changes to
Java programs to understand the ``Plastic Surgery Hypothesis'' underlying
certain types of program repair, an important but orthogonal
concern. Kim et al.~\cite{kim2013} manually analyze 
changes to Java programs to inform an automated repair
technique and show that doing so results in higher-quality repairs; 
we attempt to validate their templates heuristically on
a different dataset, and take their results as
motivation for understanding human repairs (as it may result in better patches).
Long and Rinard learn probabilistic models from bug fixes to C
code~\cite{Long2016}. Our analysis is significantly less precise and does not
inform a new repair technique.  Rather, it serves as a starting
point for research investigating the repair of bugs in Java code.

@@

\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre-fix and
post-fix version. If the count is different between pre- and post-fix
version, we consider the COB pattern found.

@@

Table~\ref{tab:freqpattern} estimates the prevelance of PAR templates in the
Java dataset. The most common pattern we observed is CBC; and the least common
pattern is AOB. If we assume that these patterns never appear together, they
cover 14.78\% of buggy files. Kim et al. mentioned that the patterns that they
use cover almost 30\% of real patches~\cite{kim2013}. Although we only detect 8
out of 10 bug fixing patterns, it is unlikely the final two will cover the
difference. Moreover, our detection is imprecise, and likely represents an upper
bound;  This suggests that common bug fixing patterns might not generalize
beyond an initial dataset, and that more work remains to identify a sufficiently
expressive set of mutations for high-quality automated repair of Java.

\section{Statement-level mutations}
\begin{table*}
	\centering
	\resizebox{\textwidth}{100pt}{
		\begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|} \hline
			&Assert&Break&Continue&Do& For&If&Label&Return&Case&Switch& Synch&Throw&Try&TypeDecl&While
			\\ \hline

@@

@@

Some automatic program repair approaches mutate source code using predefined
mutation operators such as addition, deletion and replacement. In order to
generate patches that make more sense to human developers, it would be
beneficial to analyze how developers mutate source code to fix bugs.

\vspace{1ex}\textbf{Approach.}
Because direct diffs are difficult to identify, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another across
this dataset.  To do this, for each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.
%   For each statement type (as tagged by the Boa infrastructure
% for Java), we process the results as follows: if number of occurrences of one
% statement type decreased post-fix and the number of another type increased, we
% say that the first statement type was replaced by the second statement type for
% that file. 
Note that this analysis doesn't distinguish the
replacement of the same statement kind, since we are counting the amount of
appearances of each statement kind.

We follow a similar approach to approximate a count of deletions and
insertions. 
For each bug fixing revision $r$ and each statement kind $k$, we count how many
statements of kind $k$ are there in revision $r$ and $r-1$ ($r-1$ means the
revision before revision $r$), and then compare the two numbers.

\vspace{1ex}
\noindent\textbf{Potential replacements.} Table~\ref{tab:likeliness} shows
the replacement likelihood for our dataset 
(each cell shows the percent of the time that the statement in the \emph{row}
was replaced by a statement of the type in the \emph{column}). For example, if
the cell in the DoStatement row (row 5) and AssertStatement column (column 2)
shows 0.81, meaning that Do statements were
replaced by Assert statements 0.81\% of the times.

Additional results (raw numbers not shown) show that the most common replacement
is replacing Return statement with If statement (appearing in 30,489 files). The
second most common replacement is replacing an If statement with a Return
statement, which had 28,536 appearances. 
By contrast, the least common replacement was an Assert
statement replacing a Type Declaration statement, which we did not observe. 
The second least common replacement was a Do statement for a Type
Declaration or a Label with a Type; we observed these once each.

The most common
\emph{replacer} statement kind (the statement kind that most commonly replaces
others) is the \texttt{If Statement} (101,366 appearances). The least
common \emph{replacer} is the \texttt{Type Declaration Statement} with 447
appearances.
The most common \emph{replacee} was the
\texttt{Return Statement} with 111,938 appearances; and the least common
\emph{replacee} was the \texttt{Type Declaration}, with 399 appearances.

\vspace{1ex}
\noindent \textbf{Potential deletions and insertions} From Figure~\ref{fig:StmtType}, we
can see that, expression, if, return, for and try catch statements are added
most often. Meanwhile, the same types of statements are also deleted most
often. This finding indicates that most bugs were fixed by changing the control
flow of the program.

\section{Discussion}

There exist threats to the validity of our study. Since we are using a DSL
provided by Boa to implement our analysis, the correctness of our analysis
depends on our programs as well as Boa itself. To mitigate the risk of errors in
implementation, we open source our implementation on
Github~\footnote{\url{https://github.com/}}.\todo{FIX THIS NOW.}  Some threats
to our study arise from limitations of Boa. For example, we rely on Boa to
identify bug fixing revisions; however, precisely accomplishing this is an open
problem. Another big limitation is that Boa does not provide an easy mechanism
to identify precise, statement-level diffs between revisions. Thus, our template
matching and analysis of code changes (by counting each statement kind) could be
inaccurate. However, our results should provide upper bounds.

We have characterized bug fixing changes in Java.
The findings of our study provide a set of useful insights for automatic
program repair tools in Java. It suggests that patterns proposed by the
state-of-the-art in generate and validate bug report for Java are not enough to
cover the entirety of bug fixes in Java. Adding more patterns is one solution,
but may not be a permanent one. Our finding on Java bug fixes characteristic
suggests that mutation based program repair may need to add new fields or
methods to fix bugs in Java and achieve a human-like written patch. A deeper
look suggests that it can also improve its search for a patch by having the knowledge of how
likely it is to replace a faulty statement by the different kinds of statements
available. Overall, our findings motivate additional study of the problem of
repair in Java, as assumptions that underlie approaches that target C are
unlikely to translate directly to this domain. 
% demonstrate that our knowledge of bug fixing
% properties in Java is still lacking and therefore more research on it are
% warranted.

%   We studied the applicability of the state-of-the-art in
% mutation operators to this dataset, 
% showing that 8 out of 10 bug fixing patterns in PAR cover only 15\% of bug
% fixes in this dataset.  Finally, investigated statement level mutations and
% summarized the likelihood of statement insertion, deletion, and replacement.
% These results suggest

% This study is directly applicable to automatic software repair. Most common
% tools for automatic software repair~\cite{kim2013,legoues2012} first localize a
% defect and then randomly select edits to 
% find a faulty statement using several different fault localization
% techniques~\cite{fry2010}, and once they have that faulty statement, they apply
% different kinds of edits to it, usually at random. Which makes it equally likely
% to replace an statement for another without looking at the kinds of statements
% being replaced. With this study, we are able to make a more accurate guess based
% on the phenomenon we observed from Table~\ref{tab:likeliness} and
% Figure~\ref{fig:StmtType}.

% For example, if we are fixing a Java program, and we have a faulty statement
% that happens to be a For Statement. Then we can look for the most likely
% statements to replace a For Statement (row 6): If statement (22.89\%
% likeliness), Return statement (21.08\%), and a While statement (13.79\% of
% likeliness to replace a For Statement). We also know that there is a 7.9\% of
% cases in which human programmers deleted this kind of statement, and in 10.6\%
% of the cases, a For statement was added to patch a bug. This way we provide
% guidelines to the automatic bug repair approaches to consider this on their
% effort to find a patch.
